Full Title

### 2019-10-05 fracGC/p.tex
story for fractionals; aiming for PLDI

@@

>>>>

\begin{itemize}

\item Quantum computing hot.

\item Unitary evolution; many algorithms promote classical reversible
functions into unitary gates (examples, including some from Quipper)

\item Common case: given a permutation between finite sets, find a
  reversible program that implements it. Answer depends on the
  available reversible primitives.

\item Example, say we want to transpose a matrix, i.e., write a
  permutation of type \verb|A*B <-> B*A| where \verb|A| and \verb|B|
  are the dimensions. Wikipedia example
  (\url{https://en.wikipedia.org/wiki/In-place_matrix_transposition}):
\begin{verbatim}
11 12 13 14 
21 22 23 24 
\end{verbatim}
  transposed to
\begin{verbatim}
11 21
12 22
13 23
14 24
\end{verbatim}
  Here \verb|A| is \verb|Nat x Nat| and \verb|B| is
  \verb|Nat x Nat x Nat x Nat| and the input and output matrices are:
\begin{verbatim}
M   = (11 , 21) , (12 , 22) , (13, 23) , (14 , 24)

trM = (11 , 12 , 13 , 14) , (21 , 22 , 23 , 23) 
\end{verbatim}
Say we are given a language like Pi that is sound and complete with
respect to permutations on finite types, we would write the
permutation like so.
\begin{verbatim}
WRITE PERMUTATION
\end{verbatim}
This code does not use additional space, i.e., it performs the matrix
transpose in constant space, i.e., performs an in-place matrix
transposition. It is well know that with additional space, one can
write more efficient matrix transpositions (explain and citations).

\item So many reversible/quantum circuits allow the introduction of
  so-called ancilla wires, which serve as scratch space (goes back to
  Toffoli). The problem is that there is a discipline for using
  ancilla bits: they need to be created in a known state and can only
  be gc'ed when they are returned to the same state (explain).

\item No language knows how to automatically guarantee this condition;
  left to the programmer (Quipper), or some (incomplete?) dynamic
  checks (Ricercar).

\item We solve this using **fractional types**.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Extra for now}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fractional Pi}

Adding variables to Pi first?
